### 2019-05-02 thesis/img/plot.tex

\providecommand{\myrootdir}{..}
\documentclass[tikz]{standalone}
\onlyifstandalone{
\usepackage{pgfplots}
% This small hack is required to set the search path correctly
% Just provide the current directory this file is in relatively to the root dir
\pgfplotsset{table/search path={./img/}}
}
\begin{document}
\begin{tikzpicture}
 \begin{axis}[
   xlabel={input $i$},
   ylabel={result},
   grid=major,
   legend pos=north west,
  ]
  \addlegendentry{$i+2$}
  \addplot[blue,mark=x,thick] table[y index=1] {\myrootdir/data/plot-mult.dat};
  \addlegendentry{$2 \cdot i$}
  \addplot[red,mark=x,thick] table[y index=2] {\myrootdir/data/plot-mult.dat};
  \addlegendentry{$-1^{i}$}
  \addplot[green,mark=x,thick] table[y index=3] {\myrootdir/data/plot-mult.dat};
\end{axis}
\end{tikzpicture}
\end{document}