### 2019-05-02 thesis/mini/abstract.tex

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

\begin{document}

\chapter*{\myAbstractTitle}

It is just a small template, that is hopefully helpful for writing your thesis.
First, it gives a short introduction to its usage and the personalizations.
Finally, it presents several examples of frequently used pieces of layout.

\end{document}