### 2019-05-02 thesis/mini/acknowledgments.tex

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

% Just delete this file, if you don't want to acknowledge anything.
% The corresponding page disappears automatically from all output documents

\begin{document}
\thispagestyle{empty}

\vspace*{2cm}

\begin{center}
{\usekomafont{section} \myAcknowTitle}
\end{center}

\vspace{1cm}
\noindent
If you want, you can use this section to thank your supervisors, your universities or just anybody else who has helped you writing your thesis.
But for this template, I want to abuse this section for a little request.
I cannot and do not want to track who is using this template, but I am curious how many students have actually used it.
So in case you wish to give something back to the author of this template, please consider writing an email to \href{mailto:danke@hutzelmann.net}{\nolinkurl{danke@hutzelmann.net}}.
A simple \enquote{Thanks} is already enough, but if you want to provide more details, please add the name of your study program, the name of your thesis and tell me, what you liked and disliked the most about this template.
\todo{Write this email \smiley{}}
Thank you very much for your help!

\end{document}