### 2019-05-02 thesis/zyx/TUMinfo/1-cover.tex

\providecommand{\myrootdir}{../..}
\documentclass[\myrootdir/main.tex]{subfiles}

\begin{document}

\begin{titlepage}
\centering

\includegraphics[height=2.5cm]{\logotum}

\vspace{1cm}

{\LARGE \textsc{\ifthenelse{\equal{\mylanguage}{de}}{Fakultät für Informatik}{Department of Informatics}}}\\[0.8em]
{\Large \textsc{\ifthenelse{\equal{\mylanguage}{de}}{Technische Universität München}{Technical University of Munich}}}

\vspace{5em}
{\Large \myThesisType}
\vspace{5em}

\begin{spacing}{1.1}
\Huge \bfseries \myTitle
\end{spacing}

\vspace{1.5em}

{\Large \myAuthor}

\vfill

\includegraphics[height=2.5cm]{\logoinf}

\cleardoublepage
\end{titlepage}
\end{document}