### 2019-06-09 paper/paper.tex
several example extraction programs, description of what is extractable, readme updates

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}

% TODO remove XD
\definecolor{ballblue}{rgb}{0.13, 0.67, 0.8}
\usepackage[color=ballblue]{todonotes}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-9999-9/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Analyzing Buildlogs Using Programming by Example}

\author{Carolin Brandt}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Brandt, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  ...
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%\begin{CCSXML}
%<ccs2012>
% <concept>
%  <concept_id>10010520.10010553.10010562</concept_id>
%  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%  <concept_significance>500</concept_significance>
% </concept>
% <concept>
%  <concept_id>10010520.10010575.10010755</concept_id>
%  <concept_desc>Computer systems organization~Redundancy</concept_desc>
%  <concept_significance>300</concept_significance>
% </concept>
% <concept>
%  <concept_id>10010520.10010553.10010554</concept_id>
%  <concept_desc>Computer systems organization~Robotics</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
% <concept>
%  <concept_id>10003033.10003083.10003095</concept_id>
%  <concept_desc>Networks~Network reliability</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
%</ccs2012>
%\end{CCSXML}
%
%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{ci, buildlogs, programming by example}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

shortening fixing time for broken builds important to increase dev productivity,

to detect underlying failure huge buildlog has to be read, dev has to find relevant information between a lot of lines unimportant for the task,

extracting the dev-relevant information from logs and present only them in a structured way could be improvement on dev productivity,

traditional: regexes built for exact information extraction

hard to maintain, regex understanding effort, don't generalize / have to be redone when a bulidlog changes / language or build tool is switched,

... what we do: collected a diverse set of buildlogs from travis ci, designed a metamodel of the contained (extractable?) information, labeled the collected buildlogs, built a tool which uses the PROSE library and can extract information from buildlogs based on examples, evaluated tool/learning ability/accuracy on created dataset, results ...

explicit: contributions

\section{Related Work}

\subsection{Continuous Integration}
several researchers done this and that

The software engineering research community started looking into buildlogs as more as continuous integration becomes a vital part of a modern software engineering process.
Seo et al.~\cite{seo2014programmers} found that few error types such as dependency issues and type mismatches are most prominent cause of build failures at google and that most failures are resolved within two builds.
Vassalo et al.~\cite{vassallo2017a-tale} found that the reasons for CI builds to fail are very different at open source and industrial projects using java. 
Beller et al.~\cite{beller2017oops} 
3 log analysis papers

\subsection{Tool}
improving time developers take to understand buildlogs, vasallo by summarizing relevant information in java/maven buildlogs and augmenting with links to related stack overflow  posts, amar et al. (by removing lines that appear both in passing and failing buildlogs and therefore) reducing the lines needed to be inspected, information retrieval techniques to identify most important lines for investigation and categorize the build failure cause

bart paper, icse paper???

relation to our work: we could support this research by enabling easier data extraction from the analyzed buildlogs, most focus on single/few languages with our tool they could more easily cover multiple ones

\subsection{Information Extraction and Retrieval}
flashextract paper, icse paper???

Le et al. developed FlashExtract as part of the PROSE Framework, tool to synthesize programs based on a few given examples, extract substrings from semi-structured texts, users can extract multiple fields and structure them with hierarchy and sequence, uses inductive synthesis algorithm to synthesize intended program, eliminates the need for the user to understand structure of entire document or to be able to code

we are applying their library to the domain of buildlogs, using programming by example to take away the need to tediously develop and maintain the regular expressions, traditionally used in information extraction

\section{Method}
\subsection{Data collection}
describe process of collecting buildlog files, sampling
querying ghtorrent and travis api with (insert fancy name for data collection tool) for x buildlogs of each status of the top-watched repositories on github that also use travis ci,
\subsection{Meta-Model}
through manual examination of y of the buildlogs we collected designed a metamodell  of the contained/extractable/useful/interesting information,

... explain metamodel + figure ...
\subsection{Our Tool}
... interaction and usage and output of tool, maybe integration possibilities ...

... necessary or interesting implementation details ...

\section{Results}
... achieved accuracy, samples needed for same accuracy as related tools (travistorrent), aging of learned programs with newer data ...

\section{Analysis}
... our tool works well for defined cases, limitations of tool/prose/regex usage in general, every x weeks extraction has to be rewritten to keep accuracy, well improved if only y new examples have to be added ...

\section{Conclusion and Future Work}
... repeat everything and results ...

... future: make avaliable to other researchers, integrate into travis, common extend evaluation and data collection, use dataset to evaluate against existing approaches ...

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.