### 2019-07-07 evaluation/Evaluation.Rmd
add missing logfiles, add evaluation plotting code

---
title: "Evaluation of Experiment Data"
output: html_notebook
---
# Evaluating our experiment data

## Preparing the data

```{r}
# Load the packages required to read XML files.
library("XML")
library("methods")
library("graphics")

setwd("/Users/Laci/Documents/Delft/master-thesis/pbe-extraction-buildlogs/results")

# Convert the input xml file to a data frame.
oldw <- getOption("warn")
options(warn = -1)

xml <- xmlTreeParse(file = "android-failure-with-dependencies_ManualSelection.xml")

nanosecsToTime <- function(ticks=numeric){
  sec <- ticks / 1e7
  return(as.POSIXct(sec, origin="1970-01-01", tz="UTC"))
}

data <- data.frame(ExampleCount = numeric(), TestOutput = character(), DesiredTestOutput = character(), Accuracy = numeric(), Successful = logical(), LearningDuration = as.POSIXct(character()), ApplicationDuration = as.POSIXct(character()), stringsAsFactors=FALSE)
#print(str(data))
#str(data)
#results <- xmlRoot(xml)#[['Results']]
results <- xml[['doc']]$children$EvaluationResultOfString[['Results']]
#print(results)
for (i in seq_along(1:length(results))) {
  item <- results[[i]]
  exampleCount <- length(item[['key']][[1]][['Examples']])
  test <- item[['value']][[1]][[1]]
  testOutput <- xmlValue(test[['Output']][[1]])
  desiredTestOutput <-  xmlValue(test[['DesiredOutput']][[1]])
  accuracy <- 3 # TODO
  successful <-  as.logical(xmlValue(test[['Successful']][[1]]))
  learningDuration <-  nanosecsToTime(as.numeric(xmlValue(test[['LearningDuration']][[1]])))
  #print(as.numeric(xmlValue(test[['ApplicationDuration']][[1]])))
  applicationDuration <-  nanosecsToTime(as.numeric(xmlValue(test[['ApplicationDuration']][[1]])))
  #rs <- getNodeSet(xml, '//Results/item')
  #print(typeof(sapply(rs, function(x) x[['key']][[1]][['Examples']])[[1]]))#[[0]][['Examples']])

  data = rbind(data, data.frame(ExampleCount=exampleCount, TestOutput=testOutput, DesiredTestOutput=desiredTestOutput, Accuracy=accuracy, Successful=successful, LearningDuration=learningDuration, ApplicationDuration=applicationDuration, stringsAsFactors=FALSE))
}
#print(results_set)
print(data)
  options(warn = oldw)
```

## Calculating Accuracy
```{r}
library("stringdist")

escapeStringAsNotRegex <- function(x=character()){
  return(gsub("([.|()\\^{}+$*?]|\\[|\\])", "\\\\\\1", x))
}

# print(data$TestOutput)
# print('')
# print(data$DesiredTestOutput)
for (row in 1:nrow(data)) {
  testOutput <- data[row, "TestOutput"]
  desiredTestOutput <-data[row, "DesiredTestOutput"]
  successful <- grepl(escapeStringAsNotRegex(testOutput), desiredTestOutput)
  data[row, "Successful"] <- as.logical(successful)
  # print(nchar(testOutput))
  # print(nchar(desiredTestOutput))
  # print(testOutput)
  # print(desiredTestOutput)
  accuracy <- nchar(testOutput)/nchar(desiredTestOutput)
  accuracy2 <- stringsim(testOutput, desiredTestOutput)
  print(accuracy)
  print(accuracy2)
  data[row, "Accuracy"] <- accuracy
}
print(data)
```

## Plotting the data
```{r}
plot_data <- str(data[c("ExampleCount", "Accuracy", "LearningDuration", "ApplicationDuration")])
print(str(plot_data))
plot(plot_data)
```