### 2019-10-03 thesis/txt/6-conclusion.tex

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

\begin{document}

\chapter{Conclusion}
\label{conclusion}

\section{Future Work}

\end{document}