### 2019-10-08 thesis/mini/abstract.tex
literature research about CI and benefits

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

\begin{document}

\chapter*{\myAbstractTitle}

It is just a small template, that is hopefully helpful for writing your thesis.
First, it gives a short introduction to its usage and the personalizations.
Finally, it presents several examples of frequently used pieces of layout.

\secfunc{section function - \textbackslash secfunc} \\
\mention{things to mention / reference - \textbackslash mention} \\
\plan{what to write here - \textbackslash plan} \\
\bp{actual bullet points - \textbackslash bp} \\
\draft{final text drafty - \textbackslash draft} \\
\review{final text review ready - \textbackslash review} \\
\todo{ToDo - \textbackslash todo} \\

\end{document}