### 2019-10-09 thesis/txt/4-tool.tex
writing on models chapter, putting pictures into figures, update label names

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

\begin{document}

\chapter{Our Information Extraction Tool}
\label{sec:implementation}
describe implemantation of Tool
\begin{figure}[h]
	\centering
	\includegraphics[page=7, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Our tool unifying different information extraction techniques for build logs}
	\label{fig:tool}
\end{figure}

\section{Common Interface}
\begin{lstlisting}
Usage: ruby run-extraction.rb -a analyze -t <technique: ir, pbe, random> -e <example_set> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a evaluate -t <technique: ir, pbe, keyword, random> -e <example_set> -s <selection_technique> -l <step_count_for_learning> -c <test_count>
       ruby run-extraction.rb -a analyze -t keyword -k <keyword, keyword, ...> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a analyze -t regex -r <regex to match extraction> -p <path_to_file_to_analyze>

Specific options:
    -a, --action ACTION              Either run an extraction for a example set ('analyze') or run the whole evaluation of it ('evaluate')
    -t, --technique TECHNIQUE        The technique used for creating the extraction program (pbe, ir, keyword, random)
    -e, --example-set EXAMPLE_SET    The filename of the example set to use
    -p, --path PATH                  The path to the file to be analyzed relative to the 'tool/samples' folder
    -s, --selection SELECTION        The example sequence selection technique to use for evaluation (chronological, random, manual (= like defined in file))
    -l, --learning-step-count COUNT  How many steps with increasing example set size to do during evaluation
    -c, --test-count COUNT           How many test files to evaluate the generated program in each learning step of the evaluation
    -v, --verbose                    Print additional interesting output apart from only the extraction output
    -k, --keywords X,Y,Z             Keywords too filter lines for during keyword search based extraction
    -r, --regex REGEX                Regex to match on build log file content in regex based extraction

Common options:
    -h, --help                       Show this message
\end{lstlisting}
unification tool, reference CLI somehow nicely?,

ruby tool, unifies different syntax command line interfaces of the other tools, easy to switch between different techniques (NOT TOO LONG)

\section{PROSE Program Synthesis}
implementation details :)

C\#, based on PROSE Text Extraction DSL, parsing example set files, feeding learning session \& applying learned regex program, can do single string extraction and also sequence extraction \& showing differentiating examples (although not in use)

\section{IR Similarity}
impl details,

R, text2vec, tokenized vectors, impl of preprocessing?

don't repeat from IE models section!

\section{Keyword search}
impl details,

R, simple!

\end{document}