### 2019-10-24 thesis/txt/6-conclusion.tex
writing a lot in chapter 2, various figures, data analysis and figures in thesis

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

\begin{document}

\chapter{Conclusion}
\label{sec:conclusion}
answer research questions (again)

\section{Future Work}
\label{sec:fw}
\begin{itemize}
  \item more techniques (which kinds of techniques? express in model! e.g.\ diffs might not work?), same evaluation, ir automated keyword search/expansion
  \item other works with data set: what else could this be used for?
  \item refine / improve techniques with the results automatically obtainable for evaluation with data set
    \begin{itemize}
      \item pbe: ranking adjustments, or even new dsl components?
      \item ir: meta-parameter optimizations
    \end{itemize}
  \item socio-technical: how do developers interact with buildlogs?
\end{itemize}

\end{document}