### 2019-11-07 thesis/mini/acknowledgments.tex
tud logo

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

% Just delete this file, if you don't want to acknowledge anything.
% The corresponding page disappears automatically from all output documents

\begin{document}
\thispagestyle{empty}

\vspace*{2cm}

\begin{center}
{\usekomafont{section} \myAcknowTitle}
\end{center}

\vspace{1cm}
\noindent
%I want to thank Moritz Beller for: support when coming to delft, great support and detailed feedback. invaluable contribution to this work and the large amount of things I learned about reserach in these last five months
% annibale for his neverending input on study ideas, his expertise on information retrieval
% prof. pretschner for his feedback and critical questions on the whether my evaluation is actually meaningful
% prof. knapp?
% Taico for his relentless support, upcheering in the emotional valleys of throwing everything out of the window and listening to my enthusiastic detailed descriptions of ?
% Michi for his calming input and critical reflection of my working ethics
% my new founld colleagues from SERG (xavier, enrique, ayoushi, pouria, luis, joseph), sparking joy in lunch and coffee breaks, good breaks for my focussed mind
% the fourth floor group in welcoming me in a new country as I was one of them and showing me that even when you go to a completely new place you can find awesome friends there as well

% If you want, you can use this section to thank your supervisors, your universities or just anybody else who has helped you writing your thesis.
% But for this template, I want to abuse this section for a little request.
% I cannot and do not want to track who is using this template, but I am curious how many students have actually used it.
% So in case you wish to give something back to the author of this template, please consider writing an email to \href{mailto:danke@hutzelmann.net}{\nolinkurl{danke@hutzelmann.net}}.
% A simple \enquote{Thanks} is already enough, but if you want to provide more details, please add the name of your study program, the name of your thesis and tell me, what you liked and disliked the most about this template.
% \todo{Write this email \smiley{}}
% Thank you very much for your help!

\end{document}