### 2019-11-28 thesis/txt/3-data-set.tex

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

\begin{document}

\chapter{LogChunks Data Set}
\label{sec:data-set}
\includegraphics[page=5, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
This chapter describes the creation of the \emph{LogChunks} data set.
\emph{LogChunks} encompasses about 800 build logs produced by a wide variety of programming languages.
For each build log we manually labeled which substring describes why the build failed.
Four our evaluation of the chunk retrieval techniques PBE, CTS and KWS we added search keywords and structural categories to each of the labeled substrings.

First, we present related data sets and show how \emph{LogChunks} differs.
Second, this chapter describes our log collection process and details the data schema of \emph{LogChunks}.
Further, it presents the labeling process and how we validated the obtained data points.
In our first validation study, a second person also labeled a sample of the build logs.
For the second validation study, we sent out mails to the original project developers to determine whether the substring we labeled really describes the reason the build failed.

\section{Motivation}
In/output examples (I/O examples) are used to configure two of the information retrieval techniques we are evaluating for this thesis, PROSE regular expression program synthesis by example and text similarity.
One example set always contains build log in/output examples from the same software repository or project and for one specific build log information. They are all from one repository because we investigate information retrieval techniques which are always configured to the scope of one particular software repository or project.
Our intuition was that the ability of PROSE to be able to successfully learn a regex program depended on the structural uniformity of the provided in/output examples.
The regular expressions need a consistent pattern in the build log string at the borders or around the textual representation of a build log information to match on for the retrieval.
To be able to quantify this intuition in our evaluation we assigned \emph{structural categories} to each of the examples within an example set.

\section{Related Data Sets}

\paragraph{TravisTorrent}
The \emph{TravisTorrent} data set~\cite{beller2017travistorrent} collects a broad range of metadata about builds on Travis CI\@.
It combines data accessible through the public Travis CI API~\cite{travisci2019apidoc}, related data from GHTorrent~\cite{gousios2013ghtorrent}, the corresponding git repository and data obtained through analysis of the build logs.
The data obtained through build logs contains the names of failing test cases, similar to the reason the build failed in \emph{LogChunks}.
However, these values are obtained through a manually developed parser, which only supports specific Ruby test runners and Java Maven or JUnit logs.
\emph{LogChunks} provides manually labeled data points on the description of the build failure reason for a much wider selection of programming languages.

\paragraph{Travis CI Build Log Data Set}
Loriot et al.~\cite{loriot2019dataset, loriot2019styler} collected a large amount of Travis CI build logs from 130 github repositories to analyze their use of the Checkstyle plugin.
They selected Maven repsitories that included the Checkstyle plugin and also used Travis CI\@.
Their data set only provides the plain build logs, whereas \emph{LogChunks} provides manually analyzed build logs from Travis CI\@.

\section{Log Collection}
In the following we describe how we select the repositories, builds and Logs for \emph{LogChunks}.
To collect the build logs we built the  \texttt{GHTorrentParser},\texttt{LogCollector} and \texttt{TravisRequester} using Ruby.

\paragraph{Repository Sampling}
First, we determine a set of repositories to query logs from.
Our \texttt{GHTorrentParser} queries the \emph{GHTorrent}~\cite{gousios2013ghtorrent} data set for the most popular languages on GitHub~\cite{github2019website}.
It then retrieves the most popular repositories for a given language.
We define \emph{Popularity} as the number of watches.
The \texttt{TravisRequester}, our tool querying the Travis API~\cite{travisci2019apidoc}, can then check for a given repository whether it uses Travis CI\@.

For \emph{LogChunks} we queried GHTorrent from 01/04/2018 for the three most popular repositories of each of the 30 most popular languages.

\paragraph{Build Sampling}
The \texttt{LogCollector} uses the \texttt{TravisRequester} to obtain the newest builds for a given repository.
Wwe use a stratified sampling approach: \texttt{TravisRequester} saves the obtained builds in buckets according to their status.
We encountered the following statuses during our data collection:
\begin{multicols}{3}
\begin{itemize}
	\item created
	\item started
	\item cancelled
	\item passed
	\item errored
	\item failed
\end{itemize}
\end{multicols}
The user of \texttt{TravisRequester} configures how many builds should be checked and how many builds per status should be saved.

To sample the builds for \emph{LogChunks} we let \texttt{TravisRequester} check up to a 1000 builds per repository and kept ten for each status.
We discarded all statuses except one, namely ``failed''~\cite{travis2009buildstatus}, to make the most use of Travis' existing categorization of build status.

\paragraph{Log Sampling}
For each build the \texttt{TravisRequester} then selects a log to download.
Travis CI attributes logs to \emph{jobs}.
A single build can consist of multiple jobs, e.g.\ building the same code version and executing tests in various different testing environments.
A failed build can have successful job executions, as just one failed job leads to the whole build being marked as failed.
\texttt{TravisRequester} queries each build for the first job, which has the same state.
For the selected jobs, the tool queries the Travis API V3 over HTTP to obtains the corresponding build log.

We manually inspected the collected build logs and had to discard logs from three repositories.
One had only one failed build, two others had empty build logs on Travis CI\@.
In total we collected 797 logs from 80 repositories.

\section{Data Schema}
\label{sec:data-schema}
For each repository, \emph{LogChunks} has a list of \emph{examples}.
Each example consists of:
\begin{itemize}
	\item \textbf{Input:} the relative path to the input build log.
	\item \textbf{BuildFailureReason:} the substring of the log describing the reason the build failed.
	\item \textbf{Keywords:} keywords a developer would use to search for the BuildFailureReason.
	\item \textbf{Category:} a categorization of the structural representation of the BuildFailureReason within the build log.
				The category is relative to the other examples for the same repository.
\end{itemize}
Following, this section describes the BuildFailureReason, Keywords and Category in more detail.

\paragraph{BuildFailureReason}
The \emph{BuildFailureReason} is the substring of the build logs that describes why the build failed.
This could be the failing test case, the description of a failed linter rule or a compiler error.
The \emph{BuildFailureReason} is one continuous string cut from the build log.
If there are multiple errors leading for the build to fail, the substring contains the first appearing continuous error descriptions.
\emph{Continuous} means that no lines reporting of normal build behavior are interrupting the error descriptions.
Wherever possible, it does \emph{not} include the log statements describing \emph{that} the build failed, but the description of \emph{why} it failed.
For a few logs we were unable to define the section detailing why the build failed, e.g.\ because this information was logged in another log file.
In these instances the \emph{BuildFailureReason} contains the lines describing that the build failed.

\paragraph{Keywords}
The \emph{Keywords} contain a list of one to three keywords appearing within the \emph{BuildFailureReason} or in the area around it in the build log.
We aim to select keywords a developer would use to search for the \emph{BuildFailureReason} when inspecting the build log.

\paragraph{Category}
For each repository, we assign \emph{structural categories} to the examples.
The structural category compares how the \emph{BuildFailureReasons} are represented within the build logs.
Build tools highlight their error messages with markings, e.g.\ starting each line with ``\texttt{ERROR}'', surrounding lines filled with special characters or additional empty log lines.
Two examples fall into the same structural categories if they are surrounded by similar markings.
For most cases, two BuildFailureReason examples which fall into one category are outputted either within the same build phase or by the same build tool.
For each repository, the structural categories are represented as integers, starting at 0 and increased with the next appearing category in chronological build order.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\textwidth, clip]{img/big-study/categories-dataset.pdf}
	\caption{Distribution of the categories in LogChunks}
	\label{fig:categorycount-examplecount-dataset}
\end{figure}

\section{Labeling Process}
\label{sec:labeling-process}

\paragraph{BuildFailureReason}
For each repository, the labeler skimmed through the build logs and tried to identify the first occurrence of a description why the build failed.
They copied out the first continuous description as the BuildFailureReason.
They preserved whitespace and special characters, as they might be crucial to detect the targeted substring.
To support exact learning of regular expressions identifying the labeled substrings the labeler aimed to start and end the labeled substring at consistent locations around the fault description.

\paragraph{Keywords}
We presented the BuildFailureReason substring and ten lines above and below to the labeler.
Their task was to note down three strings they would put into a document search function to find this failure description.
The string should appear in or around the BuildFailureReason substring and is case-sensitive.
There are no special limitations on the string itself, especially spaces are also allowed.
Regular Expressions are not allowed.

\paragraph{Category}
To label the \emph{structural categories} we again presented the BuildFailureReason and the surrounding context to the labeler for all logs from a repository.
We asked them to assign numerical categories according to whether the BuildFailureReason substring had the same structural representation, i.e.\ the same surrounding or identifying characters.
The labeler should start the categories with 0 and increase as new ones appear.
For reproducibility we presented the logs in chronological build order.

\section{Validation}
We validate our collected data points in two different ways.
A different labeler performed a second pass of labeling the build failure reason, keywords and categories on a subset of the data.
In addition, we sent out a survey to the developers, whose commits triggered the builds within our data set.
We asked them whether our retrieval of the log part describing the reason the build failed was correct.
This sections describes these two validation studies.

\subsection{Inter-Rater Reliability Study}
To evaluate the validity of our labeled data points we perform a second labeling of a sample from \emph{LogChunks}.

\paragraph{Method}
A second labeler processed 30 random generated logs for each data point in \emph{LogChunks}.
We followed the same labeling process as described in Section~\ref{labeling-process}.
For the build failure reason and the keywords, we presented 30 randomly sampled build logs from distinct repositories in \emph{LogChunks}.
As the structural categories are always relative to the other logs from the same repository we randomly sampled 3 repositories and presented all 10 examples within them to the second labeler.

\paragraph{Results}
For the first labeled data point, namely the substring describing why the build failed, the two labelers exactly agreed in six cases.
In 15 cases the second labeler selected more lines than the first one, in five there was partial overlap and in 4 they completely disagreed.

Regarding the keywords the two labelers completely agreed in nine cases and completely disagreed in two cases.
In nine cases there was partial overlap in the keywords of the two labelers, in two the first labeler selected additional keywords compared to the second, while the second labeler labeled additional keywords in seven cases.

When classifying the labeled substrings into structural categories, the two labelers agreed in 26 cases and disagreed in four cases.

\paragraph{Discussion}
The results of this validation study show that there is overlap in the data from the two labelers, however also a high variation.
We mainly attribute this to not specific enough instructions to the second labeler.
There were implicit, inconsistent assumptions both labelers created during their work.
In the following we describe the inconsistent assumptions between both labelers for each data point and the implications on our description of the data classes.

For the first data class, the reason the build failed, it was ambiguous whether the labeled substring should contain the information \emph{that} the build failed.
This concerns statements like \texttt{The build exited with 1}.
The second labeler included such statements, while the first one only focussed on the log parts describing \emph{why} the build failed, e.g.\ the name of the failing test case.

While labeling the keywords a developer would use to search for the log part describing why the build failed, the first labeler allowed arbitrary, case-sensitive strings.
In contrast to that, the second labeler focussed on actual \emph{words}, delimited by spaces or special characters, while ignoring capitalization.
A second difference was that the first labeler was presented with all substrings from a repository, yielding more general keywords than the second labeler.
Their keywords were in part specific to the failure itself and therefore would not have caught a log part describing a different reason for the build to fail in the test build log.

For the structural categories this validation study showed a high overlap.
In our instructions to the second labeler we did not emphasize the \emph{structural} aspect enough.
They sorted into categories along \emph{why} the build failed, putting failing tests from different test runners into the same category even though the failing tests were presented differently in the log.

Our main learning from this study is that adequately communicating all decisions and assumptions on how data is labeled is important and difficult.
We reviewed the misunderstandings and incorporated more thorough descriptions of our data classes in Section~\ref{sec:data-schema}.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/dev-mail.png}
	\caption{An example of the mails we sent out to developers for validation of our labeled log part}
	\label{fig:dev-mail}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/dev-survey.png}
	\caption{Survey for the validation with developers}
	\label{fig:dev-survey}
\end{figure}

\subsection{Developer Survey}

\paragraph{Introduction}
For \emph{LogChunks} we analyzed around 800 build logs from different repositories and tried to extract the part of the log which describes why the respective build failed.
As we are not involved in the development of any of the open source projects within our data set we could only rely on our previous experience with different build logs and systems.
We only looked into the logs and did not check the related configurations, so it is well possible that we extracted parts that do describe errors but the respective step failing is ignored by the configuration and the build failed for another reason.

The person who probably knows best why a build failed is the one committing the changes which triggered the build.
If the build was e.g.\ part of a pull request then developer likely looked at it and tried to fix it so the pull request could be accepted.
In this section, we describe how we validate the extractions of the build failure reason \emph{LogChunks} describes.
\paragraph{Method}
Using the Travis API, for every build log in the data set we looked up the corresponding build and the committer information.
We grouped all commits triggered by one developer and sent out an email to each of them, asking whether the log part selected during our labeling was indeed describing the reason the build failed.Figure~\ref{fig:dev-mail} shows one of the mails sent out.
The email included links to the corresponding commits, build overview and log file.
We asked the receivers to fill out a short survey in case our extraction was not correct.
Look at Figure~\ref{fig:dev-survey} to get an impression of the survey.
In the survey we presented the selected log part and asked the developer to paste in the log part actually describing the failure reason or describe why we were wrong in their own words.
As some of the extractions we labeled are many lines long, we trimmed all down to 10 lines to keep the mail readable.

\begin{figure}[htbp]
	\centering
	\begin{minipage}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth, clip]{img/dev-mails/answers-received-mails.pdf}
		\caption{Number of mails answered, unanswered and not delivered}
		\label{fig:mails-answers-received-mails}
	\end{minipage}\hfill
	\begin{minipage}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth, clip]{img/dev-mails/extraction-correct.pdf}
		\caption{Label correctness as validated by developers}
		\label{fig:mails-extraction-correct}
	\end{minipage}
\end{figure}
\begin{figure}[htbp]
	\centering
	\begin{minipage}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth, clip]{img/dev-mails/answers-received-builds.pdf}
		\caption{Proportions of logs answered, undelivered and unanswered}
		\label{fig:mails-answers-received-builds}
	\end{minipage}\hfill
	\begin{minipage}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth, clip]{img/dev-mails/logs-per-mail.pdf}
		\caption{Average number of logs per sent out mail}
		\label{fig:mails-logs-per-mail}
	\end{minipage}
\end{figure}

\paragraph{Results}
In total we sent out mails to 246 developers, asking about 3.2 build logs on average.
32 of these mails could not be delivered, e.g.\ because they were addressed at noreply mail addresses.
These 32 mails related to 68 of the build logs.
We received answers from 61 developers, responding about 144 build logs.
Figure~\ref{fig:mails-answers-received-mails} and Figure~\ref{fig:mails-answers-received-builds} show the proportions of mails and logs answered about, not delivered and unanswered.

Of the 144 answers, 132 said our extraction was correct.
26 answered either ``close, but not quite correct'' or ``no, the build failed for another reason''.
We manually inspected these negative answers and found that extractions were correct after all.
This yields 12 log extractions in our example set that were not correct, shown in Figure~\ref{fig:mails-extraction-correct}.

\paragraph{Discussion}
This study highly strengthens the trust in the validity of the extracted build failure reasons in \emph{LogChunks}.
The study received answers about 18\% of the logs labelled in our build log, of which, after manual correction, 91\% said our labelled extractions were accurate.

Most of the initial ``incorrect'' answers we adjusted in them manual inspection, stated that the proposed extraction did not show the whole reason the build failed.
This is because we had to trim long labeled extractions to keep the emails readable.

One of our extractions only showed a warning and the developer propose to also include the line above, stating that warnings are treated as errors in the build.
In others that were identified as ``incorrect'' we labelled the error message of an error that later ignored and did not lead to the build failing.

\end{document}