### 2019-12-18 thesis/mini/acknowledgments.tex

\providecommand{\myrootdir}{..}
\documentclass[\myrootdir/main.tex]{subfiles}

% Just delete this file, if you don't want to acknowledge anything.
% The corresponding page disappears automatically from all output documents

\begin{document}
\thispagestyle{empty}

\vspace*{2cm}

\begin{center}
{\usekomafont{section} \myAcknowTitle}
\end{center}

\vspace{1cm}
\noindent
I would like to thank Moritz for his great support, detailed feedback and his invaluable teaching how to do proper research.
I could not have imagined a better advisor for my master's thesis.

\noindent
Furthermore, I would like to thank Annibale for his helpful feedback and teaching me which is the proper vocabulary to describe my work.

\noindent
I want to pay my special regards to Prof. Pretschner for his critical questions, leading me to continuously challenge my evaluation and the problems my work addresses.

\noindent
I would like to acknowledge Martin, Martijn and Dominik for their thorough reviews.

\noindent
I am extremely grateful for my newfound friends, the fourth floor group and my colleagues from SERG.
Thank you for welcoming me with open arms and sparking joy in my days.

\noindent
Lastly, I want to thank Michi and Taico.
Thank you for your calming input, critical reflection, your relentless support and upcheering.

%I want to thank Moritz Beller for: support when coming to delft, great support and detailed feedback. invaluable contribution to this work and the large amount of things I learned about reserach in these last five months, aufbauendes feedback
% annibale for his neverending input on study ideas, his expertise on information retrieval, telling me the proper vocabulary to use and unlimited study and plotting ideas
% prof. pretschner for his feedback and critical questions on the whether my evaluation is actually meaningful and what problems I really solve
% prof. knapp? sharp eyes with formulas and /aufbauend/ words
% Michi for his calming input and critical reflection of my working ethics / balance.
% A big thank you to Martin, Taico, Martijn and Michi for their reviews.
% my new found colleagues from SERG (xavier, enrique, ayoushi, pouria, luis, joseph), sparking joy in lunch and coffee breaks and beer evenings
% the fourth floor group in welcoming me in a new country as I was one of them and showing me that even when you go to a completely new place you can find awesome friends there as well
% Lastly, Taico for his relentless support, upcheering in the emotional valleys of throwing everything out of the window and listening to my enthusiastic detailed descriptions of ?

% If you want, you can use this section to thank your supervisors, your universities or just anybody else who has helped you writing your thesis.
% But for this template, I want to abuse this section for a little request.
% I cannot and do not want to track who is using this template, but I am curious how many students have actually used it.
% So in case you wish to give something back to the author of this template, please consider writing an email to \href{mailto:danke@hutzelmann.net}{\nolinkurl{danke@hutzelmann.net}}.
% A simple \enquote{Thanks} is already enough, but if you want to provide more details, please add the name of your study program, the name of your thesis and tell me, what you liked and disliked the most about this template.
% \todo{Write this email \smiley{}}
% Thank you very much for your help!

\end{document}