### 2020-03-03 tse-paper/abstract.tex
include loads of text and graphics from thesis, fix graphics, start adapting heading levels

Build logs are textual by-products that a software build process
creates, often as part of its Continuous Integration (CI)
pipeline. Build logs are a paramount source of information for
developers when debugging into and understanding a build
failure. Recently, attempts to partly automate this time-consuming,
purely manual activity have come up, such as rule- or
information-retrieval-based techniques.

We evaluate and compare three different techniques that aim to retrieve
specified log parts (chunks) from a build log, namely program synthesis
by example, textual similarity and search keywords.
We conduct an empirical study by comparing these techniques on the manually
labeled \emph{LogChunks} data set.
Our findings show that none of the three techniques in general outperforms
the others. We discuss under which circumstances each technique performs best
and provide a recommendation on when developers or researchers should use which
technique.

% Software builds from continuous integration produce detailed logs about the status and results of the various tools involved in the build.
% These build logs are a valuable data source for developers and researchers to inspect test results, the duration of build steps or understand the cause of a build failure.
% However, build logs are very verbose, at best semi-structured and their structure differs highly between projects.
% This makes it hard to process and analyze them.
% In this paper, we evaluate and compare three different techniques that aim to retrieve specified text chunks from a build log, namely program synthesis by example, textual similarity and search keywords.
% We conduct an empirical study by comparing these techniques on the \emph{LogChunks} data set of 797 Travis CI logs from a diverse range of projects.
% Our findings show that none of the three techniques clearly outperforms the others.
% We discuss under which circumstances each technique performs best and provide a recommendation on when developers or researchers should use which technique.