### 2020-04-02 tse-paper/0-introduction.tex
split into separate files per section

% TODO: Introduction structure
% 0) we introduce & define the problem 1) we survey which techniques exist 2) we implement (for a lack of readily available implementations) and compare the most promising ones 3) we draw conclusions for when to use which technique

% TODO: Include term predictive CI, link to FB paper and https://www-sciencedirect-com.tudelft.idm.oclc.org/science/article/pii/S0164121218301730

% TODO: What about the alternate solution, installing plugins to report precise data? Ie., maven plugins
% - Lots of effort
% - not a generic solution
% - how to analyze history where such plugins were not installed
% - unclear which information you want

% TODO It would be futile to compare our techniques with one of the existing manual regex parsers, as (a) as our literature survey showed, their scope is mainly limited to the Java ecosystem and (b) whether they would work or not would be randomly based on whether the project we select exhibits features their regular expressions were trained for. Finally, they are not really automatic

\lstset{
  morekeywords={},
	basicstyle=\ttfamily\scriptsize,
  postbreak=\mbox{\textcolor{blue}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  stringstyle=\color{Plum},
	frame=single,
  extendedchars=false,
  texcl=false,
  aboveskip=\baselineskip,
  belowskip=0pt
}

Continuous Integration (CI) has become a common practice in software
engineering~\cite{hilton2016usage}.  Many software projects use
CI~\cite{hilton2016usage,staahl2014modeling,beller2017oops} to detect
bugs early~\cite{vasilescu2015quality,duvall2007continuous}, improve
developer productivity~\cite{miller2008hundred,hilton2016usage} and
communication~\cite{downs2012ambient}.  CI builds produce logs which
report results of various sub-steps within the build.  These build logs
contain a lot of valuable information for developers and researchers---for
example, descriptions of compile errors, linter warnings or failed
tests~\cite{beller2017oops,seo2014programmers,vassallo2017a-tale}.

CI builds produce logs which report the results of the steps within the build.
The information stored in build logs can and has
already been used for a variety of applications. One prominent dataset
in the space, TravisTorrent, was featured as the MSR Data Challenge
2017~\cite{msr17challenge}. To date, scientist and practitioners alike
have used it and other, proprietary datasets to understand and
cateogrize Continuous Integration build
failures~\cite{islam2017insights}, to do research on testing
practices~\cite{orellana2017differences}, to train classifiers to
predict the build outcome and
duration~\cite{ni2017cost,bisong2017built,machalica2019predictive},
and to investigate static analysis tools in CI
builds~\cite{zampetti2017open}. Therefore, being able to efficiently
and correctly extract information from build logs is paramount to the
future of a variety of fields depending on it.

However, build logs can be verbose and large---sometimes in excess of
50 MB of ASCII text ~\cite{beller2017oops}---making them inadequate
for direct human consumption. Therefore, to support developers and
researchers in efficiently making use of the information within build
logs, we must at least semi-automatically retrieve the chunks of the
log that describe the targeted information.

There are different techniques to retrieve information chunks from CI
build logs. Beller et al. use a rule-based system of regular
expressions to analyze logs from Travis CI~\cite{beller2017oops}.
Such regular expressions are developed by looking at exemplary build
logs.  Vassallo et al.\ wrote a custom parser to gather information
for build repair hints~\cite{vassallo2018un-break}.  Recently, Amar et
al.\ reduced the number of lines for a developer to inspect by
creating a diff between logs from failed and successful
builds~\cite{amar2019mining}.

These approaches have various strengths and weaknesses: Regular
expressions are exact, but tedious and error-prone to
maintain~\cite{michael2019regexes}.  Custom parsers are powerful
though fragile in light of changes in the log structure. Diffing
between failed and successful logs can reduce the information to be
processed, but is at best semi-automatic~\cite{amar2019mining}.

To gain an impression on how researchers are using information from build logs
within their research and how they are gathering chunks from them to support
developers in understanding build outcomes, we conducted a systematic mapping
study.
Our goal was to classify how and what for researchers use build logs as a source
of data (\textbf{RQ1}).

The outcomes of our study shows that there are various attempts to retrieve
specific information from buildlogs. The resulting implementations are rarely
available and therefore rarely reused. Because of the high development overhead
of the custom parsers, many implementations are limited to small number of supported
build tools.

To address this, we implement three promising techniques for build log analysis.
They retrieve specific chunks of targeted information and are configured by
examples provided by the user.
We conduct an empirical study on the \emph{LogChunks} data set~\cite{brandt2020logchunks} to
gauge the performance of these chunk retrieval techniques and characterize
when one of the techniques should be preferred over other alternatives (\textbf{RQ2}).

\begin{simplebox}{Research Questions}
\begin{itemize}
  \item[\textbf{RQ1:}] Which build log analysis techniques exist?
  \item[\textbf{RQ2:}] Which criteria influence the suitability of a chunk
  retrieval technique for build logs?
\end{itemize}
\end{simplebox}

% At the moment there is only anecdotal evidence on the performance of these
% techniques and on when a technique should be preferred over other alternatives.
% Developers and researches currently have little support when choosing which
% technique to use for a task.

% The goal of this article is to investigate different chunk retrieval techniques
% for build logs and describe under which circumstances certain techniques can be
% recommended over others.
% We aim to characterize different chunk retrieval techniques.
% For \textbf{Research Question 1} (\textbf{RQ1}), we analyze which criteria
% influence the suitability of a chunk retrieval technique for CI build logs.

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and targeted log chunks, PBE synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole text for
  the occurrence of specific trigger words.
\end{itemize}

\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* What went wrong:
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Could not determine the dependencies of task `:app:jacocoTestDebugReport'.
> Task with path 'testDebug' not found in project ':app'.
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]

* Try:
  \end{lstlisting}  
  \caption{Example of a log chunk explaining why an Android/Gradle build failed.}
  % TODO Moritz: we can relate this to travistorrent maybe, you told me long ago that this was something your analyzer cannot extract
  \label{lst:chunk-example-1}
\end{figure}

\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* What went wrong:
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Execution failed for task ':app:connectedAndroidTest'.
> com.android.builder.testing.api.DeviceException: java.lang.RuntimeException: ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]

* Try:
  \end{lstlisting}  
  \caption{Another example of a log chunk explaining why an Android/Gradle build failed.}
  \label{lst:chunk-example-2}
\end{figure}

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition, RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}, escapeinside=//}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
StartExtraction(RegexPair("Colon/{\color{Plum}$\circ$}/Line Separator", "ALL CAPS")) in EndExtraction(RegexPair("/{\color{Plum}$\varepsilon$}/", "Line Separator/{\color{Plum}$\circ$}/Line Separator"))
  \end{lstlisting}  
  \caption{Simiplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example-1,lst:chunk-example-2}}
  \label{lst:prose-program-simplified}
\end{figure}

\lstset{
  language=,
  morekeywords={},
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
--- FAIL: TestSeparator (0.03s)
    repo\_test.go:75: expected entry to be in form of `* [link] - description.`, got '* ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}  
  \caption{Example of a log chunk. \todo{which build tool?}
  \todo{move to where structural category explanation}}
  \label{lst:chunk-example-3}
\end{figure}

The results of our empirical comparison study on the chunk retrieval techniques
show that the structural representation of the targeted chunk greatly influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example-1,lst:chunk-example-2} show two exemplary log chunks
from Android/Gradle build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed. When PBE is trained on these two examples it produces
a regular expression similar to \Cref{lst:prose-program-simplified}. The retrieval
starts after a colon followed by a line separator and before a capital letter.
It ends before two line separators, consistent with the two example chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2}.
This shows, that the characters structuring the text around the chunk are
cruicial for regular expressions to identify the targeted log chunk.
In fact, if the third
example from \Cref{lst:chunk-example-3} with a different structural
representation is added, PBE is not able to syntesize a program and returns
``\texttt{no program found}''.
We express this difference in structural representation through dividing log
chunks into structural categories. The chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2} are in the same structural
category, while the chunk from \Cref{lst:chunk-example-3} is in a different one.

% \textbf{RQ2} asks under which conditions PBE, CTS and KWS are suited to retrieve information from continuous integration build logs.
% \textbf{RQ2} is refined into sub-questions along with the criteria resulting from \textbf{RQ1} and compares their instantiations for the three techniques:
% how many training examples a technique needs to perform best (\textbf{RQ2.1}), how structurally diverse the examples can be (\textbf{RQ2.2}) and how accurate the retrieved output is (\textbf{RQ2.3}).
% To evaluate PBE, CTS and KWS we use the \emph{LogChunks} data set, which encompasses about 800 log files from 80 repositories.
% Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural representation within the log.

Our study of the three techniques on \emph{LogChunks} shows that
\begin{itemize}
  \item PBE yields very accurate results when trained with two examples from a
  single structural category.
  \item CTS shows the best average precision, though precision and recall of a
  retrieval is hard to determine from the given result.
  A small increase in the number of training examples has no noticeable influence. 
  Fewer structural categories improve precision and recall of the retrieval.
  \item KWS has the highest recall of all techniques, however much lower precision.
  It is the technique with the best recall when multiple structural categories
  are present in the training examples.
\end{itemize}

% \noindent
% \textbf{Our work contributes:}
% \begin{itemize}
%   \item A tool with three prototypical implementations of chunk retrieval techniques:
%         \begin{itemize}
%           \item program synthesis from examples using the Microsoft PROSE library (PBE),
%           \item a common information retrieval approach using text similarity (CTS), and
%           \item a keyword search approach (KWS).
%         \end{itemize}
%   \item Recommendations for the configuration of each of the investigated chunk retrieval techniques.
%   \item Guidelines on choosing a suitable chunk retrieval technique.
% \end{itemize}

We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in precision and
recall of the chunk retrieval is required.
CTS is well suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by a human.
In cases where the textual representation of the desired information in the log
is unpredictable or varies greatly, KWS seems to be the best choice. However,
its low precision---it extracts a context of multiple lines around a
finding---makes it generally unsuited for automatic on-ward processing and
instead requires a human to further inspect and interpret the output chunk.