### 2020-04-02 tse-paper/2-rw.tex

\section{Related Work}
% TODO reorder rw by build retrieval technique rather than by paper
\label{sec:rw}
We explain how researchers gather information about CI usage in software projects
through build log analysis and why the chunk retrieval techniques we investigate
simplifies their data collection.
This section moves on to past works about augmenting build logs to make it easier
for developers to inspect them.
The presented approaches are related to and can benefit from the techniques we
analyze in this article.
Further, we classify build logs as semi-structured data and differentiate our
work from system log analysis.

\subsection{Build Failures in CI}
Various researchers studied why CI builds fail and into the impact of
build failures on the development workflow.

Seo et al.~\cite{seo2014programmers} find that a small group of error
types such as dependency mismatches are the most prominent cause of
build failures at Google. In addition, they notice that most failures
are resolved within two builds. They based their analyses on sets of
build logs collected from industry partners. They develop a custom
parser to classify error messages reported by Java and C++ builds.

Rausch et al.~\cite{rausch2017empirical} analyze CI builds of open
source Java projects and find that most builds fail because of failing
tests. For most projects, over half of the failed builds follow a
previous failed build. Rausch et al.'s data shows that most failures
occur in the second half of the build runtime, which can cause long
delays in the feedback loop, especially when builds are automatically
retried upon failure.

Vassallo et al.~\cite{vassallo2017a-tale} compare open source projects
in Java to industrial ones. They determine that testing failures are
more common than compilation errors. Open source builds fail most
often because of unit tests, whereas release preparations are the
primary cause in industrial projects. Their data stems from analyzing
build logs from \emph{TravisTorrent}~\cite{beller2017travistorrent}
with regular expressions.

Ghaleb et al.~\cite{ghaleb2019studying} aim to identify noise in build
breakage data. They classify build failures from \emph{TravisTorrent}
according to whether they were caused by an environmental failure or
caused by a developer change. Their analysis starts with manual
categorization of build logs. They select keywords and strings that
identify their targeted categories and code a script to automatically
classify logs based on these keywords.

\noindent
\textbf{Supporting Log Analysis with Chunk Retrieval}
To leverage the valuable information within build logs the researchers
presented in this section create parsers and regular expression-based
programs. This task of retrieving specific chunks of text from the
build logs can be solved by the chunk retrieval techniques we compare
in this article. Our results can support researchers in choosing a
suitable technique for their data set of build logs and the chunks
they want to retrieve. By relieving them from building custom parsers
we enable them to cover a much wider range of languages and build
tools in their studies.

\subsection{Augmentation of Build Logs}
\label{sec:rw-bl-analysis}
Build logs are a valuable data source for developers to find out why
their build failed. Several researchers investigate how to support
developers in working with verbose build logs. Vassallo et
al.~\cite{vassallo2018un-break} try to shorten the time it takes
developers to understand build logs. They parse
Maven~\cite{maven2019website} build logs into a structured
representation and create hint generators.
%The hint generators leverage this structured access to the
%information within the build log to propose fixes.
For example, one of the hint generators queries stack overflow for
discussions related to why the build failed. In a qualitative study
they observed that highlighting the locality and context of an issue
is helpful to programmers.
%Their tool BART is published as a Jenkins
%Plugin~\cite{bart2019plugin}.
The chunk retrieval techniques we compare in this article can be used
to fill similar structured representations with information from build
logs. As they simplify the construction of parsers they would enable
developers and researchers to cover a wider array of build tools,
which is the main influence factor on the structure of a build log.

Amar et al.~\cite{amar2019mining} compare different approaches to
reduce the portions of a log that a developer has to inspect. Their
techniques remove lines that appear both in logs from passing and
failing builds and use a modified \emph{Term Frequency Inverse
Document Frequency} (TF-IDF) weighting to identify term vectors likely
to occur with failures. Their diff technique can be interpreted as a
chunk retrieval technique, where the targeted information is defined
by the past failures used as basis for the weighted term vectors.

Travis itself already provides build log augmentation in the form of a
basic structuring of the build logs within their log viewer using
\emph{log folds}~\cite{travis2019logfolds}. They add fold identifiers
around common commands and setup or teardown build phases and collapse
the contained lines by default.

\subsection{System Log Analysis}
\label{sec:log-analysis}
A related field of log processing is the processing of system log
files produced during runtime. A main difference between build logs
and system logs is that system logs are fundamentally structured
through events. Each line in a log file represents one event with a
set of fields: timestamp, verbosity level and raw message
content~\cite{he2017towards}. Figure \ref{lst:system-log} shows
example lines from a system log.

The first goal in analyzing system log files is ubiquitously to
separate constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}. Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters. The output of a log
parser is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}. This structured log is then
the input to various machine learning and data mining processes.
%Researchers mine patterns for operational
%profiling~\cite{nagappan2009efficiently},
%debugging~\cite{oliner2012advances}, performance analytics or anomaly
%detection~\cite{nagappan2010abstracting}. Xu et
%al.~\cite{xu2009detecting} leverage the connection of log statements
%to the source code producing them to separate messages into constant
%and variable parts more accurately.

The techniques developed for system log analysis can also be applied
to build logs. One example is comparing execution traces to reference
traces of intended behavior to detect anomalies. Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
%Classic log parsers interpret the whole log file into a sequence of
%events. A similar approach could also be applied to build logs to
%determine the sequence of executed build steps or phases.
In this article, we focus on extracting a single specified information
from the build log as a whole with chunk retrieval techniques. Chunk
retrieval techniques are used as a part of log parsing to retrieve the
values of variable parts in a log message, e.g. by using regular
expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
  keywordstyle=\color{Plum},
  escapeinside=**
}
\begin{figure*}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet Responder: Received block blk_3587508140 of size 67108 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140 terminating
  \end{lstlisting}  
  \caption{System Log excerpt. Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

\subsection{Program Synthesis by Example}
\label{sec:rw-prose}
% \section{Information Extraction and Retrieval Techniques} The
% techniques we investigate are based on existing methods of
% Programming by Example.%, information extraction and information
% retrieval. This section presents different Programming by Example
% resources surrounding the PROSE library. We explain its generic
% program synthesis algorithm and how PROSE synthesizes text
% extraction programs, the foundation of PBE\@.

Programming by Example (PBE) is a form of generalized program
synthesis that enables users to automate repetitive tasks, without
them having to explicitly codify a computer program. One has to
provide prototypical examples of the input and the desired output. A
synthesis algorithm tries to create a program (based on a regular
language) that transforms the input to the output. The minimum working
condition is for this when replayed on the examples inputs is to
provide the specified output, but, hopefully, PBE is as close as
possbile to the user's intention and thus also works on a different
and larger set of inputs. This section introduces the theoretical
foundations of the program synthesis algorithm of the \emph{PROgram
Synthesis using Examples} (PROSE) framework~\cite{prose2019webpage}.
The PROSE framework is developed by Microsoft Research. Next, this
section presents the FlashExtract DSL, which defines text extraction
tasks within PROSE and is the basis for the implementation of our
chunk retrieval technique PBE\@.

\subsubsection{FlashMeta: Inductive Program Synthesis}
The FlashMeta framework presented by Polozov and
Gulwani~\cite{polozov2015flashmeta:} is the backbone of the program
synthesis in the Microsoft PROSE framework. FlashMeta separates the
inductive synthesis algorithm from the domain specific capabilities of
the desired program by encoding the possible program space in a domain
specific language (DSL). The user specifies the desired program
behavior by providing in/output examples (I/O examples). FlashMeta
uses \emph{witness functions}, provided by the DSL, to divide the
synthesis into smaller subtasks. For each of these subtasks it
enumerates all possible programs that solve the subtask consistent
with the set of I/O examples. A program is consistent with a set of
I/O examples if, for each input example, it produces the corresponding
output~\cite{mitchell1982generalization}. The possible subprograms are
joined and stored in a \emph{version space algebra}
(VSA)~\cite{mitchell1982generalization}. This a tree structure, which
space-efficiently saves candidate programs for tasks by sharing common
subexpressions. Next, FlashMeta ranks the enumerated programs
according to which ones the user most likely intended. The DSL also
provides the ranking characteristics. From the ranked VSA, FlashMeta
can then return a ranked list of complete programs consistent with the
user's example.

% In addition to I/O examples of the intended program, the user can
% also provide examples with only input or negative input examples.
% Negative input examples should not be processed by the synthesized
% program.

% The different applications of PROSE presented in the following
% paragraphs were all eventually implemented as DSLs for the FlashMeta
% synthesis algorithm.

\subsubsection{FlashExtract: Data Extraction by Example}
Le et al.~\cite{le2014flashextract:} developed FlashExtract as a DSL
for the Microsoft PROSE framework.
% It enables a user to define text extraction programs for text,
% websites and spreadsheets by giving I/O examples.
FlashExtract's instantiation for text synthesizes extraction programs
from semi-structured text based on regular expressions. Users can
extract multiple fields and structure them with hierarchy and
sequence.
% FlashExtract synthesizes programs to extract each of the fields
% leveraging the information about hierarchical containment and
% sequentiality. It eliminates the need for the user to understand the
% entire structure of the processed document and decreases the effort
% of developing a suitable extraction program.

FlashExtract models the extraction of a single substring as a pair of
two cut positions. A position is either specified by an absolute
character index or by a pair of two regular expressions. The first
regular expression matches the substring directly before the
characterized position, the second regular expression matches the
substring directly after. A regular expression in FlashExtract is a
concatenation of tokens, e.g.\ standard character classes or string
literals frequently occurring in the input examples. Figure
\ref{lst:prose-program} shows a text extraction program synthesized by
FlashExtract. This program defines the first position as after a colon
followed by a newline character and before a piece of text with all
capital letters. It defines the second position as before two newline
characters.

Apart from automatic completion in Excel
spreadsheets~\cite{excel2019flashfill}, FlashExtract is the basis for
two other Microsoft product features: Microsoft's system log analysis
tool Azure Monitor lets users define custom log
fields~\cite{azure2019custom}. The ConvertFrom-String function in
PowerShell allows a user to specify an example template to extract
hierarchical data from a text document~\cite{powershell2019convert}.

We apply the text instantiation of FlashExtract to the domain of build
logs with our chunk retrieval technique PBE\@.

\lstset{
  language=caml,
  morekeywords={PosToEndRegion, RegexPosition, RegexPair, StartToPosRegion},
  keywordstyle=\bfseries\color{black},
  escapeinside=//
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
let s = v in let s = PosToEndRegion(s, RegexPosition(s, RegexPair("Colon/$\circ$/Line Separator", "ALL CAPS"), 1)) in StartToPosRegion(s, RegexPosition(s, RegexPair("/$\varepsilon$/", "Line Separator/{\color{blue}$\circ$}/Line Separator"), 1))
  \end{lstlisting}  
  \caption{Text extraction program synthesized by FlashExtract.}
  \label{lst:prose-program}
\end{figure}