### 2020-04-02 tse-paper/7-threats.tex

\section{Threats to Validity}
There are several threats to the validity of the conclusions of our
work.

\noindent
\textbf{Implementation}
\label{sec:prose:impl}
% TODO we need to say something about why we think that these
% limtiations do not actually harm our results
Our results depend on the implementation of the investigated chunk
retrieval techniques and the libraries we used. Our implementation of
PBE is based on the program synthesis provided by PROSE\@. The
idiosyncrazies of this framework influence our PBE results. Other
frameworks similar to PROSE might have somewhat different strengths
and weaknesses. For example, the need for examples from a single
structural category stems from the fact that PROSE cannot learn
regular expression programs with arbitrary boolean conjunctions and
disjunctions~\cite{mayer2015user}. PROSE introduced this constraint to
keep the synthesis performance reasonable. At the same time, this is
clearly a current theoretical challenge of all PBE implementations,
and we therefore attribute it to the technique itself, rather than the
specific implementation.

% TODO cite http://text2vec.org/
Our implementation of CTS is dependent on the library {\tt text2vec}
and the way it splits strings into word tokens. On the other hand,
{\tt text2vec} is the de-facto standard library to do word embeddings
in R. We intentionally chose a simple, minimally configured and tuned
approach to compare against. Tuning the text similarity
meta-parameters more to the specific use case of chunk retrieval from
build logs would yield better chunk retrieval results.

\noindent
\textbf{Data Set}
The outcomes of our comparison study are highly dependent on the build
logs from the \emph{LogChunks} data set. It only consists of build
logs from open source projects and therefore it is not clear whether
our results are generalizable to industry projects. We only collected
build logs from Travis CI, however we chose to evaluate on an
information chunk whose format is not dependent on Travis CI\@. This
is because the reason the build failed is described within the build
logs by the tools themselves and not the Travis CI environment.

\noindent
\textbf{Training Set Size}
Especially the results for CTS might be influenced by the fact that we
only trained on one to five examples. We chose this small training set
size as the training examples have to be provided per repository and
we expect a developer to not want to provide more examples than the
small numbers we evaluated on.

\noindent
\textbf{Few Samples with Many Structural Categories}
Our comparison study shows fewer measurements with many structural
categories than with one category.
This stems from the fact that we investigated the chunk retrieval
techniques on a realistic data set, which showed to often have few
structural categories within one project.