### 2020-04-02 tse-paper/8-future-work.tex

\section{Future Work}
\todo{survey build logs tools in industry}

There are various future research opportunities based on our work:
\begin{itemize}
  \item \textbf{Further Analysis of \emph{LogChunks}} We created the
  \emph{LogChunks} data set \cite{brandt2020logchunks} specifically for the comparative
  study in this paper, though it can be the basis for various further
  analyses of build log data. The keywords, for example, can be
  investigated to answer which keywords are used to search for the
  reason the build failed within build logs.
  \item \textbf{Cross-Repository Build Log Analysis} We trained and
  tested each chunk retrieval technique on examples from the same
  repository. We propose to analyze how techniques could be trained
  across repositories, building the cornerstones for build
  environment-agnostic analysis tools.
  \item \textbf{Comparison with more Chunk Retrieval Techniques} This
  paper investigates the three chunk retrieval techniques PBE, CTS and
  KWS\@. Our study design can be reused to evaluate other build log
  analysis techniques, such as the diff and information retrieval
  approach by Amar et al.~\cite{amar2019mining}.
  \item \textbf{Refinement of Retrieval Quality for each Technique} We
  investigated basic configurations of existing techniques applied to
  chunk retrieval from build logs. In a next step, each of these
  techniques could be refined to better approach the domain of build
  logs. The \emph{LogChunks} data set and our study results act as a
  baseline to benchmark such technique improvements. We propose the
  following improvements:
    \begin{itemize}
      \item \textbf{Custom Ranking and Tokens for PBE} The program
      synthesis through PROSE ranks possible programs according to
      what the user most likely intended. One could adapt the ranking
      rules provided by the FlashExtract DSL to fit common build log
      chunk retrieval tasks. FlashExtract includes special tokens when
      enumerating possible regular expressions. One could extend these
      with tokens found in build logs, such as ``-'',``='',``ERROR''
      or ``[OK''.
      \item \textbf{Meta-Parameter Optimization for CTS} Information
      retrieval techniques have various meta-parameters which can be
      optimized for the specific use
      case~\cite{panichella2016parameterizing}. We propose to further
      investigate improvements in preprocessing of the log text, in
      tokenization of the log lines into terms and in stop words
      lists.
    \end{itemize}
  \item \textbf{Usability Analysis of Chunk Retrieval Output} Our
  analysis of the output produced by the chunk retrieval focuses on
  precision and recall. We propose to investigate how useful these
  outputs are to developers through controlled experiments.
\end{itemize}