### 2020-04-22 tse-paper/abstract.tex
First feedback and review integrated

Build logs are the textual by-products that automatic software build processes
create, often as part of their Continuous Integration (CI)
pipeline.
Build logs can serve as a paramount source of
information for developers to understand and debug depdency, compilation, or
test failures.
Manually
extracting the important chunks of information in a build log, though, is akin to
finding a needle in a haystack.
Recently, practitioners and researchers have started attempts to partly
automate
this time-consuming activity, to not only ease
developers' task of finding this needle but also to enable a novel class of
automated on-ward processing applications.
In this paper, we give a systematic overview of the emerging field of build log analysis.
In a large-scale literature mapping study, we first survey and categorize the existing methods to
extract information chunks from build logs.
In the second part of the paper, we develop prototypical implementations of the three most promising
techniques we found, namely program synthesis
by example, textual similarity, and keyword search.
Finally, we evaluate the techniques in an empirical study on
the \emph{LogChunks} data set, which comprises 797 build logs 
from a wide range of different projects and programming languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths
and weaknesses.
We discuss under which circumstances each technique performs best
and provide a recommendation scheme for when developers or researchers might 
use which technique.